---
title: "Prediction Assignment Writeup"
author: "John Magcamit"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# OVERVIEW

The goal of this project is to predict the manner in which the user performed the exercise (classe). Random Forest was used to create the model for the prediction and k-fold Cross Validation was used as the resampling technique because it was faster in processing time vs. Bootstrapping. Also, bootstrapping had no positive impact on model accuracy as stated in [pml-RandomForestPerformace](https://github.com/lgreski/datasciencectacontent/blob/master/markdown/pml-randomForestPerformance.md). The Random Forest model with 10-fold cross validation produced an accuracy of 99.14%.



## Initializing Libraries and Data

```{r, message=FALSE, warning=FALSE}
library(caret)
library(dplyr)
library(parallel)
library(doParallel)

traindf <- read.csv("pml-training.csv")
testdf <- read.csv("pml-testing.csv")
```

Cleaning the Training Data by replacing blank values  to NA then removing NA columns. The first 7 columns are also removed to reduce irrelevant variables for creating the model.
```{r}
traindf[traindf==""] <- NA
traindf <- traindf[,colSums(is.na(traindf))==0]
traindf <- traindf[,-c(1:7)]
```

## Creating the model

Setting up the parallel processing to speed up building a model using random forest. (c) [pml-randomForestPerformance](https://github.com/lgreski/datasciencectacontent/blob/master/markdown/pml-randomForestPerformance.md)
```{r}
cluster <- makeCluster(detectCores()-1)
registerDoParallel(cluster)
```

```{r, cache=TRUE}
set.seed(23)
inTrain <- createDataPartition(y=traindf$classe, p=0.7, list = FALSE)
training <- traindf[inTrain,]
testing <- traindf[-inTrain,]
#setting x and y to avoid slowness of caret with model syntax
x <- training[,-53]
y <- training[,53]
fitControl <- trainControl(method="cv", number=10, allowParallel = TRUE)
fit <- train(x,y, method="rf", data=training, trControl=fitControl)
```

After finishing the training, parallel processing is closed.
```{r}
stopCluster(cluster)
registerDoSEQ()
```

Checking the models accuracy
```{r}
fit

fit$resample

confusionMatrix.train(fit)
```

Matching the columns from training data to testdf and adding the classe variable to replace the problemid column.
```{r}
testdf <- testdf[names(testdf) %in% names(training)]
testdf$classe <- as.character("")
predict(fit, testdf)
```

